---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    highlight: tango
    keep_md: yes
    theme: journal
    toc: yes
  pdf_document:
    toc: yes
---


## Loading and preprocessing the data

```{r, echo=TRUE}
setwd("C:\\Users\\Dario\\Documents\\GitHub\\RepData_PeerAssessment1")
suppressMessages(library(dplyr))
activity <- read.table(unz("repdata_data_activity.zip", "activity.csv"),
                       header = TRUE, sep = ",")
```

## What is mean total number of steps taken per day?

```{r, echo=TRUE}
stepsPerDay <- activity %>% select(date, steps) %>% group_by(date) %>% 
  summarize(mean(steps, na.rm = TRUE))
names(stepsPerDay) <- c("date", "sum")
```

1. Make a histogram of the total number of steps taken each day

```{r echo=TRUE}
hist(stepsPerDay$sum, 
     freq = TRUE,
     col = 'red', 
     xlab = 'Total Number of Steps per day',
     breaks = 10,
     main = 'Number of Steps per Day')
dev.copy(png, file = 'figures/plot1.png', height=480, width=480)
dev.off()
stepsPerDayMean <- mean(stepsPerDay$sum, na.rm = TRUE)
stepsPerDayMedian <- median(stepsPerDay$sum, na.rm = TRUE)
```

2. Calculate and report the mean and median total number of steps taken per day

Variable       |  mean       |         median
---------------|-------------|-----------------------
Steps Per Day  | `r format(round(stepsPerDayMean, 4), scientific = FALSE)` | `r format(round(stepsPerDayMedian, 4), scientific = FALSE)`

## What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval and the ave number steps

```{r echo=TRUE}
intervalPerDayMean <- activity %>% group_by(interval) %>% 
  summarize(mean(steps, na.rm = TRUE))
names(intervalPerDayMean) <- c("interval", "MeanSteps")
intervalLevels <- levels(factor(activity$interval))
par(mar = c(4.5,4.5,1,1))
plot(intervalPerDayMean$MeanSteps ~ intervalLevels, type = "l",
     ylab = "number of steps", xlab = "interval levels", col = "red")
dev.copy(png, file = 'figures/plot2.png', height=480, width=960)
dev.off()
```

2. Which 5-minute interval on average contains the maximum number of steps

```{r echo=TRUE}
max5MinMeanValue <- max(intervalPerDayMean$MeanSteps)
max5MinMeanRow <- which(intervalPerDayMean$MeanSteps == max5MinMeanValue)
max5MinMeanInterval <- intervalPerDayMean[row(intervalPerDayMean) == 
                                            max5MinMeanRow,][1,1]
```

**The (`r max5MinMeanInterval`) 5-minute interval contains 
`r format(round(max5MinMeanValue, 0), scientific = FALSE)` maximum number 
of steps in average across all days: 
**

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset

```{r, echo=TRUE}
iSteps <- which(is.na(activity[,1]) == TRUE)
jInterval <- activity[iSteps,3]
activity[iSteps, 1] <- intervalPerDayMean[,2]
```

**Total number of missing values rows in the dataset is: `r length(iSteps)`**

2. Devise a strategy for filling in all of the missing values in the dataset.

   - Strategy is to fill in the missing values with the average of all existing
     values in a particular day

``` {r echo=TRUE}
missingvals = c(NA)
meanVector <- tbl_df(aggregate(activity$steps[!activity$steps %in% missingvals],
                               list(activity$date[!activity$steps %in% missingvals]), mean))
names(meanVector) <- c("date", "steps")
meanVector$date <- as.character(meanVector$date)
newActivity <- tbl_df(activity)
newActivity$date <- as.character(newActivity$date)


mutate(newActivity, steps = ifelse(steps == NA, 
                                   meanVector$steps[newActivity$date == 
                                                      meanVector$date], steps))

newActivity$steps[newActivity$steps %in% missingvals] <- 
  meanVector$steps[newActivity$date == meanVector$date]

```

3. Create a new dataset that is equal to the original dataset but missing
   values filled in
   
   
4. 

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
stepsPerDay
```
** Si las hay **
