---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    highlight: tango
    keep_md: yes
    theme: journal
    toc: yes
  pdf_document:
    toc: yes
---


## Loading and preprocessing the data

```{r, echo=TRUE}
setwd("C:\\Users\\Dario\\Documents\\GitHub\\RepData_PeerAssessment1")
suppressMessages(library(dplyr))
suppressMessages(library(lubridate))
stringsAsFactors=FALSE
activity <- read.table(unz("repdata_data_activity.zip", "activity.csv"),
                       header = TRUE, sep = ",")
```

## What is mean total number of steps taken per day?

```{r, echo=TRUE}
stepsPerDay <- activity %>% select(date, steps) %>% group_by(date) %>% 
  summarize(mean(steps, na.rm = TRUE))
names(stepsPerDay) <- c("date", "sum")
```

1. Make a histogram of the total number of steps taken each day

```{r echo=TRUE}
hist(stepsPerDay$sum, 
     freq = TRUE,
     col = 'red', 
     xlab = 'Total Number of Steps per day',
     breaks = 10,
     main = 'Number of Steps per Day')
dev.copy(png, file = 'figures/plot1.png', height=480, width=480)
dev.off()
stepsPerDayMean <- mean(stepsPerDay$sum, na.rm = TRUE)
stepsPerDayMedian <- median(stepsPerDay$sum, na.rm = TRUE)
```

2. Calculate and report the mean and median total number of steps taken per day

Variable       |  mean       |         median
---------------|-------------|-----------------------
Steps Per Day  | `r format(round(stepsPerDayMean, 4), scientific = FALSE)` | `r format(round(stepsPerDayMedian, 4), scientific = FALSE)`

## What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval and the ave number steps

```{r echo=TRUE}
intervalPerDayMean <- activity %>% group_by(interval) %>% 
  summarize(mean(steps, na.rm = TRUE))
names(intervalPerDayMean) <- c("interval", "MeanSteps")
intervalLevels <- levels(factor(activity$interval))
par(mar = c(4.5,4.5,1,1))
plot(intervalPerDayMean$MeanSteps ~ intervalLevels, type = "l",
     ylab = "number of steps", xlab = "interval levels", col = "red")
dev.copy(png, file = 'figures/plot2.png', height=480, width=960)
dev.off()
```

2. Which 5-minute interval on average contains the maximum number of steps

```{r echo=TRUE}
max5MinMeanValue <- max(intervalPerDayMean$MeanSteps)
max5MinMeanRow <- which(intervalPerDayMean$MeanSteps == max5MinMeanValue)
max5MinMeanInterval <- intervalPerDayMean[row(intervalPerDayMean) == 
                                            max5MinMeanRow,][1,1]
```

**The (`r max5MinMeanInterval`) 5-minute interval contains 
`r format(round(max5MinMeanValue, 0), scientific = FALSE)` maximum number 
of steps in average across all days: 
**

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset

```{r, echo=TRUE}
iSteps <- which(is.na(activity[,1]) == TRUE)
jInterval <- activity[iSteps,3]
activity[iSteps, 1] <- intervalPerDayMean[,2]
```

**Total number of missing values rows in the dataset is: `r length(iSteps)`**

2. Devise a strategy for filling in all of the missing values in the dataset.

   - Strategy is to fill in the missing values with the average of all steps
     values in a particular interval

``` {r echo=TRUE}
missingvals = c(NA)

meanVector <- activity %>% group_by(interval) %>% summarize(avg_steps = mean(steps, na.rm = TRUE, na.inf = na.omit))
```

3. Create a new dataset that is equal to the original dataset but missing
   values filled in

```{r echo=TRUE}

newActivity <- tbl_df(activity)

newActivity <- cbind(newActivity, meanVector)

newActivityVector <- data.frame(newActivity$avg_steps)
newActivityVector <- cbind(newActivityVector, newActivity$date)
newActivityVector <- cbind(newActivityVector, newActivity$interval)

names(newActivityVector) <- c("steps", "date", "interval")

```
   
4. Make a histogram of the total number of steps taken each day and 
   Calculate and report the mean and median total number of steps taken per day.
   
```{r echo=TRUE}

meanf <- newActivityVector %>% group_by(date) %>% 
  summarize(avg_steps = mean(steps)) 
medianf <- newActivityVector %>% group_by(date) %>% summarize(med_steps = median(steps))
```

Variable       |  mean       |         median
---------------|-------------|-----------------------
Steps Per Day  | `r format(round(as.numeric(meanf$avg_steps), 4), scientific = FALSE)` | `r format(round(as.numeric(medianf$med_steps[1]), 4), scientific = FALSE)`

## Are there differences in activity patterns between weekdays and weekends?

Yes. It depends how reliable may be a technique for filling in NA data.

1. Create a new factor variable in the dataset with two levels - "weekday"
   and "weekend" indicating whether a given date is a weekday or weekend day.
   
```{r, echo=TRUE}
weekdays1 <- c('Mon', 'Tue', 'Wed', 'Thu', 'Fri')
activity <- activity %>% mutate(daytype = 
                                  factor((weekdays(as.POSIXlt(activity$date),
                                                   abbr = TRUE) %in% 
                                            weekdays1)+1L, levels = 1:2,
                                         labels = c("weekend", "weekday")))

```   
   
2. Make a panel plot containing a time series plot of the 5-minute interval 
   (x-axis) and the average number of steps taken, averaged across all weekday 
   days or weekend days (y-axis).
   
```{r echo=TRUE}
par(mfrow = c(2, 1))
for (i in 1:2) {
  #plot()
} 
